---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# daps

<!-- badges: start -->
<!-- badges: end -->

The goal of daps is to ...

## Installation

<!-- You can install the released version of daps from [CRAN](https://CRAN.R-project.org) with: -->
<!-- --> 
<!-- ``` r -->
<!-- install.packages("daps") -->
<!-- ``` -->

You can install the development version of `daps` from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("NikKrieger/daps")
```
## Create a `daps` object

Create a `daps-class` object and add a metastate model as well as individual variable models.

```{r load_libraries}
library(ordinal)
library(nnet)
library(tidyverse)
library(daps)
```

```{r create_daps_object}
daps1 <-
  daps() %>% 
  
  add_models(
    
    age(t) := lag(age) + 1,
    
    sbp(t) := list(
      lm(~ lag(sbp) + lag(dbp) + sex + race + age),
      lm(~ sex + race + age),
      lag(sbp),
      rnorm(1, 130, 15)
    ),
    
    dbp(t) := lm(~lag(sbp) + lag(dbp) + sex + race + age),
    dbp(t) := lm(~sex + race + age),
    
    diabetes(t) := glm(~sex + race + age + lag(sbp) + lag(sbp, 2) + lag(dbp) + lag(dbp, 2), family = binomial),
    diabetes(t) := glm(~sex + race + age + lag(sbp) + lag(dbp), family = binomial),
    diabetes(t) := glm(~sex + race + age, family = binomial),
    
    chd_risk(t) := list(
      clm(~ lag(sbp) + lag(dbp) + sex + race + age),
      multinom(~ lag(sbp) + lag(dbp) + sex + race + age)
    ),
    
    glucose(t) := lm(~sbp + dbp + diabetes + age + sex + race),
    glucose(t) := lm(~sbp + dbp + age + sex + race),
    glucose(t) := lm(~diabetes + age + sex + race),
    glucose(t) := lm(~age + sex + race),
    glucose(t) := lag(glucose)
    
  ) %>%
  
  add_metastate_model(
    # implement a "default_metastate", which is the first "state" unless the user
    # specifies
    # 
    metastate = c("state1", "state2", "state3"),
    nodes =
      list(
        c("age", "sbp", "dbp", "chd_risk", "diabetes"),
        c("age", "sbp", "dbp", "chd_risk", "diabetes"),
        c("age", "sbp", "dbp", "chd_risk", "diabetes", "glucose")
      ),
    transitions =
      list(
        ~case_when(
          lag(sbp) >= 135 | lag(dbp) >= 80                            ~ "state3",
          lag(chd_risk) == "elevated" | sex == "male" & lag(age) > 35 ~ "state2",
          TRUE                                                        ~ "state1"
        ),
        ~case_when(
          lag(sbp) >= 135 | lag(dbp) >= 80                            ~ "state3",
          TRUE                                                        ~ "state2"
        ),
        ~case_when(
          TRUE                                                        ~ "state3"
        )
      )
    # default_state = NULL
  )

daps1
```


```{r test_data}
teststatic <- 
  tribble(
    ~id, ~sex,     ~race,
    1,   "male",   "black",
    2,   "female", "white",
    3,   "male",   "black",
    4,   "male",   "white",
    6,   "female", "black"
  )

testtemporal <-
  tribble(
    ~id, ~t, ~sbp, ~dbp, ~age, ~diabetes, ~glucose, ~chd_risk,
    1,   1,  120,  65,   35,   FALSE,     NA,       "low",
    1,   2,  119,  66,   36,   FALSE,     NA,       "normal",
    1,   3,  118,  68,   50,   FALSE,     NA,       "normal",
    1,   4,  150,  100,  51,   TRUE,      NA,       "elevated",
     
    2,   1,  139,  81,   64,   FALSE,     NA,       "normal",
    2,   2,  140,  111,  66,   FALSE,     90,       "elevated",
    2,   3,  137,  85,   66,   TRUE,      100,      "elevated",
    2,   4,  155,  90,   66,   TRUE,      99,       "elevated",
     
    3,   1,  100,  40,   34,   TRUE,      NA,       "low",
    3,   2,  114,  45,   34,   FALSE,     NA,       "low",
    3,   3,  100,  50,   34,   FALSE,     NA,       "low",
    3,   4,  103,  56,   34,   FALSE,     NA,       "low",
     
    4,   1,  115,  110,  85,   FALSE,     125,      "elevated",
    4,   2,  140,  125,  86,   TRUE,      NA,       "elevated",
    4,   3,  NA,   NA,   87,   TRUE,      100,      "elevated",
    4,   4,  NA,   NA,   88,   TRUE,      NA,       "elevated",
    
    6,   1,  114,  111,  86,   FALSE,     99,       "normal"
  ) %>% 
  mutate_at("chd_risk", ordered, levels = c("low", "normal", "elevated"))
```

```{r training}
daps_trained <- daps1 %>% train(teststatic, testtemporal)

daps_trained
```


```{r simulate}
daps_trained %>% 
simulate(
  static = teststatic, 
  longitudinal = testtemporal, 
  h = NULL,
  from = "last",
  to = 10,
  impute = FALSE,
  seed = 20200123
)

daps_trained %>% 
simulate(
  static = teststatic, 
  longitudinal = testtemporal, 
  h = 1:3,
  from = "last",
  to = 3,
  impute = FALSE,
  seed = 20200123
)

daps_trained %>% 
  simulate(
    static = teststatic,
    longitudinal = testtemporal,
    h = NULL,
    from = 5,
    to = 7,
    impute = TRUE
  )

# Add in a count of patients/observations with incomplete predictions
# impute options:
#   - default: FALSE
#   - locf
#   - simulate at the preceding rows with missingness (cut it off if you have to back more than lookback_steps steps)

# daps() %>%
#   add_models() %>% 
#   add_metastate_model() %>%
#   train() %>%
#   simulate()
```

# TO DO list 2020-01-13

  - Revise the example model
    * Add a treatment variable ("none", "antihypertensives", "statins", "both")
    * Add incident outcomes within each observation year
  - Make a sufficiently deidentified data set, ideally using EHR
    * Apply for IRB to create a public data set
      - ADI, year, etc.
      - (talk to Amy Nowacki)
  - Incorporate "static" versus "dynamic" simuation. Put this switch in the initialize() function (mode = c("static", "dynamic"); add "horizon" parameter--NULL implies).
  - seed parameter in simulate(); t0 parameter in simulate
   
## Future packages

  1. Forecast accuracy - functions to join original data back up with simulated data in order to test the accuracy of the predictions
  2. Parameter uncertainty
  3. Adapting irregular temporal data to discrete time


```{r}
# nonparsnip <- 
#   lm(
#     sbp ~ lag(sbp) + lag(dbp) + sex + race + age, 
#     data = full_join(teststatic, testtemporal)
#   )
# 
# nonparsnip
# 
# parsnip <-
#   linear_reg() %>% 
#   set_engine(engine = "lm") %>% 
#   fit(
#     formula = sbp ~ lag(sbp) + lag(dbp) + sex + race + age,
#     data = full_join(teststatic, testtemporal)
#   )
# 
# parsnip
# 
# parsnip_xy <-
#   linear_reg() %>% 
#   set_engine(engine = "lm") %>% 
#   fit_xy(
#     x =
#       with(
#         full_join(teststatic, testtemporal, by = "id"),
#         cbind(lag(sbp), lag(dbp), sex, race, age)
#       ),
#     y = full_join(teststatic, testtemporal)$sbp
#   )
# 
# parsnip_xy
# 
# predict(nonparsnip, tibble(sbp = c(90, 95), dbp = c(55, 70), sex = c("male", "male"), race = c("white", "white"), age = c(50, 51)))
# 
# predict(parsnip, new_data = tibble(sbp = c(90, 95), dbp = c(55, 70), sex = c("male", "male"), race = c("white", "white"), age = c(50, 51)))
# 
# predict.model_fit(parsnip, new_data = tibble(sbp = c(90, 95), dbp = c(55, 70), sex = c("male", "male"), race = c("white", "white"), age = c(50, 51)))
```
